---
title: Workflow E4D project - working title paper erosion modeling
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'soil')`"
author:
  - given_name: Jonas
    surname: Lenz
    affiliation: "1, 2, *"
    email: jonas.lenz@iproconsult.com
    corresponding: true
  - given_name: Anette
    surname: Eltner
    affiliation: 3
    email: 
  - given_name: Lea
    surname: Epple
    affiliation: 3
    email: 
  - given_name: Anne
    surname: Bienert
    affiliation: "3"
  - given_name: Oliver
    surname: Grothum
    # authors can have multiple affiliations, which can also be used to mark deceased coauthors
    affiliation: "3"
# If you have more than one corresponding author, add them manually using the following structure (note the commas):
# Two authors: Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)
# Three authors or more: Daniel Nüst (daniel.nuest@uni-muenster.de), Josiah Carberry (j.carberry@orcid.org), and Markus Konkol (m.konkol@wwu.de)
# If the following line is uncommented, the "corresponding: true" above are ignored
#correspongdingauthors: Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)
# If an author is deceased, please mark the respective author name(s) with a dagger '†' and add a further affiliation; put the decease date in the 'address' field", see 'Nikolaus Copernicus' in template.
# If authors contributed equally, please mark the respective author names with an asterisk '*' and add a further affiliation: 'These authors contributed equally to this work.'", see template.
affiliation:
  - code: 1
    address: Institute for Geoinformatics, University of Münster, 48149 Münster, Germany
  - code: 2
    address: Psychoceramics, Wesleyan University, Middletown, CT, United States
  - code: 3
    address: University of Ferrara, Ferrara, Italy
  - code: "†"
    address: deceased, 24 May 1543
  - code: "*"
    address: These authors contributed equally to this work.
abstract: |
  The abstract goes here.
  It can also be on _multiple lines_.
bibliography: sample.bib
running:
  title: R Markdown Template for Copernicus
  author: Nüst et al.
# This section is mandatory even if you declare that no competing interests are present.
competinginterests: |
  The authors declare no competing interests.
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company. 
### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #code: |
  #  use this to add a statement when having only software code available
  #data: |
  #  use this to add a statement when having only data sets available
  codedata: |
    use this to add a statement when having data sets and software code available
  sample: |
    use this section when having geoscientific samples available
videosupplement: |
  use this section when having video supplements available
authorcontribution: |
  Daniel wrote the package. Josiah thought about poterry. Markus filled in for a second author.
disclaimer: |
  We like Copernicus.
acknowledgements: |
  Thanks to the rticles contributors!
appendix: |
  \section{Figures and tables in appendices}
  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
  \subsection{Option 1}
  If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
  They will be correctly named automatically.
  \subsection{Option 2}
  If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
  
  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
  `\appendixfigures` needs to be added in front of appendix figures
  `\appendixtables` needs to be added in front of appendix tables
  
  Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
output:
  rticles::copernicus_article: 
    highlight: NULL
    keep_tex: true
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
---

\introduction[Introduction]

Introduction text goes here.
You can change the name of the section if necessary using `\introduction[modified heading]`.

The following settings can or must be configured in the header of this file and are bespoke for Copernicus manuscripts:

- The `journal` you are submitting to using the official abbreviation. You can use the function `rticles::copernicus_journal_abbreviations(name = '...')` to search the existing journals.

- Specific sections of the manuscript:

  - `running` with `title` and `author`

  - `competinginterests`

  - `copyrightstatement` (optional)

  - `availability` (strongly recommended if any used), one of `code`, `data`, or `codedata`

  - `authorcontribution`

  - `disclaimer`

  - `acknowledgements`

See the defaults and examples from the skeleton and the official Copernicus documentation for details.

**Please note:** Per [their
guidelines](https://publications.copernicus.org/for_authors/manuscript_preparation.html),
Copernicus does not support additional \LaTeX{} packages or new \LaTeX{}
commands than those defined in their `.cls` file. This means that you cannot add any extra dependencies and a warning will be thrown if so.  
This extends to syntax highlighting of source code. Therefore this template sets
the parameter `highlight` in the YAML header to `NULL` to deactivate Pandoc
syntax highlighter. This prevent addition of external packages for highlighting
inserted by Pandoc. However, it might be desirable to have syntax highlight
available in preprints or for others reasons. Please see
`?rmarkdown::pdf_document` for available options to activate highlighting.

**Important**: Always double-check with the official manuscript preparation
guidelines at
[https://publications.copernicus.org/for_authors/manuscript_preparation.html](https://publications.copernicus.org/for_authors/manuscript_preparation.html),
especially the sections "Technical instructions for LaTeX" and "Manuscript
composition". Please contact Daniel Nüst, `daniel.nuest@uni-muenster.de`, with
any problems.


# Input Data

```{r}
library(raster)
```


## Forcing Data

Alle Daten im Fullswof input format, müssen für andere Modelle entsprechend übersetzt werden.

- DEM as xyz data file
- rainfall - constant intensity and duration as *.csv file
- soil parameters / infiltration rate, texture
- plot boundary conditions

```{r}
dgm_xyz <- read.table("models/FullSWOF/Inputs/2021-07-20_sfm_ueberstau1_nachher_trafo_Plotinneres_raster_Mean_0.10_redu_add.txt", sep = " ", header = FALSE, col.names = c("X","Y", "Z"))
dgm <- raster::rasterFromXYZ(dgm_xyz )

raster::plot(dgm)
dir.create("models/Lisem/wd")
raster::writeRaster(dgm, filename = "models/Lisem/wd/dgm.tif")

```


## Validation / calibration data

all validation data is variable in space within plot boundaries and time within experiment duration

- flow velocity data obtained from plot observation movies (particle / tracer tracking)
- time series of elevation change (DOD - richtig? und was bedeuted DOD eigentlich)
- 

# Modelierung

## Randbedingungen

- "Hydrologie" / Infiltration wird als gegeben angesehen -> wird nicht in Parameteroptimierung angefasst?
- Fokus liegt auf Hydraulik -> Abflussgeschwindigkeiten und
- Erosionsmodellierung -> DOD


## FULLSWOF

Varianten Parametrisierung

- Rauhigkeit in Parameterdatei als globalem Wert oder als Eingabedatei
- vorhergehender Modelllauf mit Infilmodel zur Berechnung von Initialbedingungen
- Fullswof Startbedingung für Wasserhöhe und geschwindigkeit

Anne legt lauffähigen fullswof Ordner ab:
 - kompilierte Exe mit Referenz auf Primärquelle + Version
 - alle Input Daten
 - Projektdatei
 
```{r run_Fullswof}
setwd("D:\\Projekte\\E4D\\paper_erosion_modeling\\models\\FullSWOF\\")
system2("FullSWOF_2D_win.exe")

```

## openLISEM

Anne legt lauffähigen openLisem Ordner im E4D Projektordner ab - analog Fullswof.
Übersetzung der Fullswof Input-Daten in LISEM Daten (maps, bzw. TIFs) sollte gecodet werden (z.B. mit R-library raster/rgdal). Aktuell über LAZ-Dateien mit proprietärer Software -> Lizenzprobleme und Geschichte.


## LISEM - Bastians Version

Anne legt lauffähigen Lisem Ordner im E4D Projektordner ab - analog Fullswof.
Checken, wie das Script (aktuell per "Pfeil" gestartet) von Kommandozeile gestartet wird

Version 0.2.4 https://sourceforge.net/projects/lisem/files/LISEM%20%28beta%20release%20new%20software%29/Portable/lisem_0.2.4.zip/download



```{r install_lisem}
if(!dir.exists("models/Lisem/bin"))
{
  download.file("https://downloads.sourceforge.net/project/lisem/LISEM%20%28beta%20release%20new%20software%29/Portable/lisem_0.2.4.zip", destfile = "models/Lisem/lisem0.2.4.zip")
  unzip("models/Lisem/lisem0.2.4.zip", exdir = "models/Lisem/")
  unlink("models/Lisem/lisem0.2.4.zip")
}

```


```{r run_LISEM}
## get instalation path to LISEM

system2("D:\\Projekte\\E4D\\paper_erosion_modeling\\models\\Lisem\\bin\\lisem_app.exe",
'-x -r "D:\\Projekte\\E4D\\paper_erosion_modeling\\models\\Lisem\\Lisem_skript\\rainfall_sim_200721.script" -d "D:\\Projekte\\E4D\\paper_erosion_modeling\\models\\Lisem\\wd\\"', stdout = FALSE
)

```

```{r}
setwd("D:\\Projekte\\E4D\\paper_erosion_modeling\\models\\Lisem\\wd\\")
lisem_results <- list.files()

velx <- raster::stack(lisem_results[grepl("velx", lisem_results)])
vely <- raster::stack(lisem_results[grepl("vely", lisem_results)])
dem <- raster::stack(lisem_results[grepl("dem", lisem_results)])
infil <- raster::stack(lisem_results[grepl("infil", lisem_results)])
wh <- raster::stack(lisem_results[grepl("wh", lisem_results)])
q <- raster::stack(lisem_results[grepl("q", lisem_results)])

rain_total <- raster::raster("rain_total.tif")
```

```{r check_waterbalance}
raster::plot(rain_total)
unique(raster::values(rain_total))
rain <- sum(raster::values(rain_total))
rain/100*1000
#unique(raster::values(rain_total))
```

```{r get_outflow}

q_steps <- list()

for(z in names(q))
{
  raster::plot(q[[z]], main = z)
  
  q_steps[[z]] <- sum(q[[z]][29,]) * 5
}
plot(unlist(q_steps))
```

```{r}
for(z in names(velx))
{
  raster::plot(velx[[z]], main = z)
}
```

```{r}
for(z in names(wh))
{
  raster::plot(wh[[z]], main = z)
}
```
## Aktuell erkannte Herausforderungen
 
 - alles in Code umsetzen
 - Numerische Stabilität -> Abbruchkriterium für Fullswof über Laufzeit der Modellierung
 - Lauffähigkeit des Workflows erstmal mit 10cm Auflösung erarbeiten, dann feiner auflösen (Laufzeit der Modelle)

# Auswertung Modellergebnisse

Zuerst:

- Auslesen der von Modellen relevanten, auf Platte geschriebenen Dateien
- Abgleich der Ausgabewerte mit Kalibrierungsdaten (z.B. stimmt an Punkt X zum Zeitpunkt Y die Fließgeschwindigkeit überein?)
- Formalisierung dieses Abgleichs in Objektivitätsfunktion ("Fitting target")

Dann:

- Auswahl Parameteroptimierungsalgorithmus -> Monte Carlo nicht unbedingt effizient
- Sensitivitäts